\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{tikz}
\usepackage{float}
\usepackage{listings}
\usepackage{color}
\usepackage{caption}
\usepackage[acronym]{glossaries}
\usepackage[nottoc]{tocbibind}
\usepackage[cache=false]{minted}
\usemintedstyle{default}
\graphicspath{{./images/}}
\newminted{cpp}{frame=lines,framerule=2pt}

\tikzstyle{bag} = [align=center]

\lstset{frame=tb,
  language=Haskell,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=2,
  stepnumber=1,
  escapechar=!
}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\title{%
      Combinatorial Problem Solving \\
      Final Project - Box Wrapping
}
\author{%
  Juan Pablo Royo Sales \\
  \small{Universitat Politècnica de Catalunya}
}
\date\today

\pagestyle{fancy}
\fancyhf{}
\fancyhead[C]{}
\fancyhead[R]{Juan Pablo Royo Sales - UPC MIRI}
\fancyhead[L]{CPS - Final Project}
\fancyfoot[L,C]{}
\fancyfoot[R]{Page \thepage{} of \pageref{LastPage}}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays

\newacronym{cp}{CP}{Constraint Programming}
\newacronym{bsp}{BSP}{Best Success Principle}
\newacronym{ffp}{FFP}{Fail First Principle}
\newacronym{bab}{BAB}{Branch and Bound}

\begin{document}

\maketitle

\section{Constraint Programming}

\subsection{Predefinitions}
Values given by the state the of the program.

\begin{itemize}
  \item $1 \leq W \leq 11$ maximum width paper roll
  \item $L$ maximum length paper roll. By definition is infinite
  \item $N = \{1, \dots, n\}$ total number of Boxes
  \item $w_b$ Width of the Box $b, b \in N$
  \item $h_b$ Height of the Box $b, b \in N$
\end{itemize}

\subsection{Formal Definition}
Given the statement of the problem proposed, we are going to define the following variables for our \acrfull{cp} solution.

\subsubsection{Variables}
\begin{itemize}
  \item $l$ is the length of paper needed for the current configuration
  \item $x_b^{tl}$: Top \textbf{left} corner $x$. Means the box $b$ is on the top left $x$-coordinate.
  \item $y_b^{tl}$: Top \textbf{left} corner $y$. Means the box $b$ is on the top left $y$-coordinate.
  \item $x_b^{br}$: Bottom \textbf{right} corner $x$. Means the box $b$ is on the bottom right $x$-coordinate.
  \item $y_b^{br}$: Bottom \textbf{right} corner $y$. Means the box $b$ is on the bottom right $y$-coordinate.
  \item $r_b$: Boolean Array that indicates if the Box $b$ is rotated or not.
\end{itemize}

As we can see this variables are obvious from the statement problem in section \textbf{2.2 Output Format} Paragraph 2 and 3.

\subsubsection{Domains}
\begin{itemize}
  \item $l = \{1,\dots,n\}$
  \item $x_b^{tl} = \{0,\dots,W-1\}, b \in N$
  \item $y_b^{tl} = \{0,\dots,L-1\}, b \in N$
  \item $x_b^{br} = \{0,\dots,W-1\}, b \in N$
  \item $y_b^{br} = \{0,\dots,L-1\}, b \in N$
  \item $r_b = \{0,1\}, b \in N$
\end{itemize}


\subsubsection{Constraints}

\begin{enumerate}
  \item $x_b^{tl} \leq x_b^{br}\ \forall b \in N$: Each top left $x$ coordinate must be less or equal than the bottom right $x$-coordinate in each Box.
  \item $y_b^{tl} \leq y_b^{br}\ \forall b \in N$: Each top left $y$-coordinate must be less or equal than the bottom right $y$-coordinate in each Box.
  \item\label{c:4} $x_b^{br} \leq W\ \forall b \in N$: There is no bottom right $x$-coordinate that is longer that the maximum width of paper roll $W$.
\item\label{c:5} $x_b^{br} = x_b^{tl} + (w_b - 1)\ \forall b \in N, r_b = 0$: If the Box is not rotated the amount of $x$ coordinates that occupies it is the width of the box.
\item\label{c:6} $y_b^{br} = y_b^{tl} + (h_b - 1)\ \forall b \in N, r_b = 0$: If the Box is not rotate the amount of $y$-coordinates that occupies it is the height of the box.
\item $x_b^{br} = x_b^{tl} + (h_b - 1)\ \forall b \in N, r_b = 1$: Same as~\ref{c:5} but the amount of $x$-coordinates that occupies is the height because it is rotated.
  \item $y_b^{br} = y_b^{tl} + (w_b - 1)\ \forall b \in N, r_b = 1$: Same as~\ref{c:6} but the amount of $y$-coordinates that occupies is the width because it is rotated.
  \item\label{c:7} $x_{b_i}^{br} + 1 \leq x_{b_j}^{tl} \lor y_{b_i}^{br} + 1 \leq y_{b_j}^{tl} \lor x_{b_j}^{br} + 1 \leq x_{b_i}^{tl} \lor y_{b_j}^{br} + 1 \leq y_{b_i}^{tl}, \ \forall b_i, b_j \in N, i \neq j$: Not overlap boxes.
  \item $x_b^{tl} = 0,\ b = 1$: Break column symmetry. At least first box should be on the most left half of the roll width.
   \item $y_b^{tl} = 0,\ b = 1$: Break row symmetry. At least first box should be on the top half of the roll length.
  \item $l = \argmax_{y_b^{br}} {b \in N}$: Length of the roll at the end cannot be larger than the biggest bottom right $y$-coordinate.
\end{enumerate}

\subsection{Optimization and Improvements}
\subsubsection{Constraint Optimization}
Regarding constraint optimization I am implemented the \textbf{length} optimization of the roll paper which is the final goal of the problem: Putting all boxes such as we are minimizing the used length. This is done in the \mintinline{cpp}{constrain} method during \acrfull{bab}.

\begin{listing}[H]
  \inputminted[firstline=153, lastline=156, linenos, breaklines]{cpp}{../src/space/box_space.cc}
  \caption{Branch and Bound Constraint Optimization}
  \label{lst:constrain}
\end{listing}

\subsubsection{Propagation Level}
Regarding propagation level all the constrains can be set with specific propagation level. In my case I select DOMAIN propagation level.

\begin{listing}[H]
  \inputminted[firstline=178, lastline=180, linenos, breaklines]{cpp}{../src/space/box_space.cc}
  \caption{Branch and Bound Constraint Optimization}
  \label{lst:prop}
\end{listing}

\subsubsection{Variable - Value Selection}
The branch and bound strategy it is being implemented only on 3 Array variables of the model: $x_tl$, $y_tl$ and $r_b$. This is because the other 2 arrays which are the bottom rights, are calculated based on the constraints stated in the previous section and not need to do a value or var selection for that.

In all cases Variable selection is \textbf{NONE} because I am controlling selection of variable ordering the boxes before started the program in decrease way according to its area.

I have implemented the following strategies for the value selection:

\begin{listing}[H]
  \inputminted[firstline=129, lastline=132, linenos, breaklines]{cpp}{../src/space/box_space.cc}
  \caption{Branch and Bound on $x_b^{tl}$ - Variable Selection}
\end{listing}

\begin{itemize}
\item \textbf{$x_b^{tl}$ Value Selection}: In this case I am trying to fit as much as possible horizontal first as we have seen in the Queens example problem.

  \begin{listing}[H]
  \inputminted[firstline=42, lastline=73, linenos, breaklines]{cpp}{../src/space/box_space.cc}
  \caption{Branch and Bound on $x_b^{tl}$ - Variable Selection}
\end{listing}


\item\label{item:var_y_b_tl} \textbf{$y_b^{tl}$ Value Selection}: In the case of $y$ top left coordinate value selection since it is depend on $x_tl$ by constrain I am just selecting the minimum.

\item \textbf{$r_b^{tl}$ Value Selection}: In the case of Rotation I am trying to test first with Rotation if possible, to do a fail fast technique.

\end{itemize}




\end{document}

