\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

\title{Lecture 02/17/2020 \\
Multidimensional Data Structures
}
\author{Juan Pablo Royo Sales}
\date\today

\pagestyle{fancy}
\fancyhf{}
\fancyhead[C]{}
\fancyhead[R]{Juan Pablo Royo Sales - UPC MIRI}
\fancyhead[L]{ADS - Lecture}
\fancyfoot[L,C]{}
\fancyfoot[R]{Page \thepage{} of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\begin{document}

\maketitle

\section{Multidimensional Data Structures}

MDS can be a lot of things, see slides.

But we can divided into:

- Points
- Spatial

What we are going to study is \textbf{Point Muiltidimensional data structures}

We identify a point in $K$-dimensional space with key $x = (x_0,x_1,\dots,x_{k-1})$

We will assume that each $x_i$ belongs to $Di = [0,1]$

\subsection{Associative Retrieval}
This kind of DS is not for doing exact search, but for doing more complexs searches.
When the search involves more than one attribute is call associative.

Some examples:

\begin{itemize}
  \item Partial Match Queries: All points that matches the attributes that i care not others.
  \item Orthogonal Range Queries: What points fall in some region. Because of that ranged queries.
  \item Nearest Neighbor Queries: Closest points to some specific point.

\end{itemize}


See slides for formal defiinitions.

\section{Data Structures}

\subsection{Inverted Files - Projections}
I am going to have an Order list for every attribute and in that list i am going to have every point. (See slides)

The list contain the points in the order according to the attribute of that list.

How do i perform an exact search?

The file is order so the complexity is $O(\log{}n)$ if i dont allow repetitions.

If i allow repetitions is $O(k\log{}n)$

For Partial match or Orthogonal Ranged Queries the cost is

$O(k\log{}n + R)$ R = Results or Number of points that match or are in the query.

\subsection{Grid Files}
It is a generalization of Hash Table.

In the 2-dimensional case it is a Matrix in which in each cell we have a list of points. See notes.

This kind of data structure is space-driven in contrast with Trees that are data-driven.

It is good for region search but we can have a lot of space empty.

Cost = See Grid Files notes.

Exact search = $O(b)$



\end{document}
